---
title: "Assignment_01"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

# Code By: Xan Lamoreux
# Advanced Machine Learning


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import Keras Library and load imdb dataset
num_words = 10000 (To keep the top 10000 most frequent words)
```{r}
library(keras)
imdb <- dataset_imdb(num_words = 10000)
```

## Get the list of reviews for train and test data and labels
Labels 0 means negative review
Labels 1 means positive review
```{r}
train_data <- imdb$train$x
train_labels <- imdb$train$y
test_data <- imdb$test$x
test_labels <- imdb$test$y
```

Check one of data set value and its label
```{r}
str(train_data[[11]])
train_labels[[11]]
```

Maximum index number

```{r}
max(sapply(train_data, max))
```


Decoding Review
```{r}
word_index <- dataset_imdb_word_index()
word_index[11]

reverse_word_index <- names(word_index)
reverse_word_index[11]

names(reverse_word_index) <- word_index

decoded_review <- sapply(train_data[[11]], function(index) {
  word <- if (index >= 3) reverse_word_index[[as.character(index - 3)]]
  if (!is.null(word)) word else "?"
})

cat(decoded_review)
```


## Preparing the Data
Use one-hot-encoding technique to turn list into vectors of 0's and 1's
```{r}
vectorize_sequences <- function(sequences, dimension = 10000) {
  results <- matrix(0, nrow = length(sequences), ncol = dimension)
  for (i in 1:length(sequences))
    results[i, sequences[[i]]] <- 1
  results
}

x_train <- vectorize_sequences(train_data)
x_test <- vectorize_sequences(test_data)

str(x_train[1,])

y_train <- as.numeric(train_labels)
y_test <- as.numeric(test_labels)
```

## Creating a Validation Set
```{r}
val_indices <- 1:10000

x_val <- x_train[val_indices,]
partial_x_train <- x_train[-val_indices,]
y_val <- y_train[val_indices]
partial_y_train <- y_train[-val_indices]

```


## Build the network
Create the sequential model
```{r}
model <- keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")
```

Compiling the model
```{r}
model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

```



## Train the model
```{r}
history <- model %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)

str(history)

plot(history)

```



## Creating model from scratch and test it
```{r}
model <- keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")

model %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

model %>% fit(
  x_train,
  y_train,
  epochs = 4,
  batch_size = 512)

results <- model %>% evaluate(x_test, y_test)

results
```

## Check predictions on new data set
```{r}
model %>% predict(x_test[1:10,])
```



## Task # 01.
## You used two hidden layers. Try using one or three hidden layers, and see how doing so affects validation and test accuracy.

```{r three_hidden_layers}
# USING 3 HIDDEN LAYERS
model_01 <- keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")


model_01 %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

history_01 <- model_01 %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)

str(history_01)

plot(history_01)

results_01 <- model_01 %>% evaluate(x_test, y_test)

results_01

history_01
```



## TASK # 02.
## Try using layers with more hidden units or fewer hidden units: 32 units, 64 units, and so on.

```{r 32_hidden_units}
# USING 32 HIDDEN UNITS
model_02 <- keras_model_sequential() %>%
  layer_dense(units = 32, activation = "relu", input_shape = c(10000)) %>%
  layer_dense(units = 32, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")


model_02 %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

history_02 <- model_02 %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)

str(history_02)

plot(history_02)

results_02 <- model_02 %>% evaluate(x_test, y_test)

results_02

history_02
```




## TASK # 03.
## Try using the mse loss function instead of binary_crossentropy.


```{r mse_loss_function}
# USING MSE LOSS FUNCTION
model_03 <- keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")


model_03 %>% compile(
  optimizer = "rmsprop",
  loss = "mse",
  metrics = c("accuracy")
)

history_03 <- model_03 %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)

str(history_03)

plot(history_03)

results_03 <- model_03 %>% evaluate(x_test, y_test)

results_03

history_03

```



## TASK # 04.
## 4. Try using the tanh activation (an activation that was popular in the early days of neural networks) instead of relu.


```{r tanh_activation_function}
# USING TANH ACTIVATION FUNCTION
model_04 <- keras_model_sequential() %>%
  layer_dense(units = 16, activation = "tanh", input_shape = c(10000)) %>%
  layer_dense(units = 16, activation = "tanh") %>%
  layer_dense(units = 1, activation = "sigmoid")


model_04 %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

history_04 <- model_04 %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)

str(history_04)

plot(history_04)

results_04 <- model_04 %>% evaluate(x_test, y_test)

results_04

history_04
```



## TASK # 05.
## 5. Use any technique we studied in class, and these include regularization, dropout, etc., to get your model to perform better on validation.


```{r adding_dropout}
# USING DROP OUT

model_05 <- keras_model_sequential() %>%
  layer_dense(units = 16, activation = "relu", input_shape = c(10000)) %>%
  layer_dropout(rate = 0.5) %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dropout(rate = 0.5) %>%
  layer_dense(units = 1, activation = "sigmoid")


model_05 %>% compile(
  optimizer = "rmsprop",
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)

history_05 <- model_05 %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 20,
  batch_size = 512,
  validation_data = list(x_val, y_val)
)

str(history_05)

plot(history_05)

results_05 <- model_05 %>% evaluate(x_test, y_test)

results_05

history_05

```



